---
title: "p8105_hw6_sy3352"
author: "Su Yan"
date: "2025-11-23"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(ggplot2)
set.seed(1)
library(p8105.datasets)
data("weather_df")
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1
```{r}
homicides_df = read_csv("data/homicide-data.csv") |> 
  mutate(location = sprintf("%s, %s", city, state),
         solved = case_when(
           disposition == "Closed by arrest" ~ 1, TRUE ~ 0),
         victim_age = as.numeric(victim_age)
         ) |> 
  filter(!location %in% c("Dallas, TX", "Phoenix, AZ", 
                          "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("Black", "White"))
```

```{r}
baltimore_df = homicides_df |> 
  filter(location == "Baltimore, MD")

baltimore_fit = baltimore_df |> 
  glm(solved ~ victim_age + victim_race + victim_sex, data = _, 
      family = binomial()) 

baltimore_results = baltimore_fit |> 
  broom::tidy(conf.int = TRUE, exponentiate = TRUE) |> 
  filter(term == "victim_sexMale") |>
  select(term, adjusted_OR = estimate, conf.low, conf.high, p.value) |> 
  knitr::kable(digits = 3)
baltimore_results
```

```{r}
by_city = homicides_df |> 
  group_by(location) |> 
  nest() |> 
  mutate(glm_model = map(data, ~glm(solved ~ victim_age + victim_race + 
                                      victim_sex, data = .x, family = binomial()
  )),
  results = map(glm_model, ~broom::tidy(.x, conf.int = TRUE, 
                                        exponentiate = TRUE))) |> 
  select(location, results) |> 
  unnest(results) |> 
  filter(term == "victim_sexMale") |>
  select(location, adjusted_OR = estimate, conf.low, conf.high, p.value) 
```

```{r}
by_city_plot = by_city |> 
  ggplot(aes(x = reorder(location, adjusted_OR), y = adjusted_OR)) + 
  geom_point(color = "deeppink", size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "City", y = "Estimated OR of Male vs Female",
       title = "Comparison for Solving Homicides by Victim Sex of Each City")
by_city_plot
```

## Problem 2
```{r}
boot_weather = weather_df |> 
  select(tmax, tmin, prcp) |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
         results = map(models, broom::tidy),
         glances = map(models, broom::glance)) |> 
  select(-strap, -models) |> 
  unnest(glances, results)

boot_results = boot_weather |> 
  select(.id, r.squared, term, estimate) |> 
  pivot_wider(names_from = term, values_from = estimate) |> 
  mutate(ratio = tmin/prcp)

boot_r_plot = boot_results |> 
  ggplot(aes(x = r.squared)) +
  geom_histogram()
boot_r_plot

boot_ratio_plot = boot_results |> 
  ggplot(aes(x = ratio)) +
  geom_histogram()
boot_ratio_plot

boot_ci = boot_results |> 
  summarize(
    r.squared_lower = quantile(r.squared, 0.025), 
    r.squared_upper = quantile(r.squared, 0.975),
    ratio_lower = quantile(ratio, 0.025), 
    ratio_upper = quantile(ratio, 0.975)) |> 
  knitr::kable(digits = 3)
boot_ci
```

## Problem 3
```{r}
bwt_df = read_csv("data/birthweight.csv") |> 
  drop_na() |> 
  rename(
    head_cm = bhead, length_cm = blength, mother_wt_deliver = delwt, 
    income_hundreds = fincome, father_race = frace, gestation_week = gaweeks,
    mother_age_mensrche = menarche, mother_height = mheight, 
    mother_age_delivery = momage, mother_race = mrace, previous_lbw_n = pnumlbw,
    previous_small_ges_n = pnumsga, mother_prepreg_bmi = ppbmi, 
    mother_prepreg_wt = ppwt, cig_smoked_n = smoken, mother_wt_gain = wtgain
  ) |> 
  mutate(
    babysex = factor(babysex, levels = c(1,2), labels = c("male", "female")),
    father_race = factor(father_race),
    mother_race = factor(mother_race),
    malform = factor(malform, levels = c(0,1), labels = c("absent", "present"))
  ) 
```

```{r}
proposed_model = bwt_df |> 
  lm(bwt ~ babysex + head_cm + length_cm + mother_wt_deliver + income_hundreds +
       father_race + mother_race + mother_height + mother_age_delivery +
       mother_prepreg_bmi + cig_smoked_n + mother_wt_gain + gestation_week, 
     data = _)
summary(proposed_model)

bwt_df_predict = bwt_df |> 
  modelr::add_predictions(proposed_model) |> 
  modelr::add_residuals(proposed_model)

bwt_predict_pt = bwt_df_predict |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "hotpink") +
  labs(x = "Fitted values", y = "Residuals")
bwt_predict_pt
```
I evaluated model assumptions using a residuals vs fitted plot generated with `add_predictions()` and `add_residuals()`. The plot shows a roughly elliptical cloud of points centered around zero with no strong pattern or evidence of nonlinearity. The variance of residuals is approximately constant across fitted values. These diagnostics suggest that the linear model assumptions are reasonably satisfied.

```{r}
model_main_effects = bwt_df |> 
  lm(bwt ~ length_cm + gestation_week, data = _)
summary(model_main_effects)

model_interaction = bwt_df |> 
  lm(bwt ~ (head_cm + length_cm + babysex)^3, data = _)
summary(model_interaction)

cv_df = modelr::crossv_mc(bwt_df, 50) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df_fit = cv_df |> 
  mutate(
    main_effect_mod = map(train, \(df) lm(bwt ~ length_cm + gestation_week, data = df)),
    interact_mod = map(train, \(df) lm(bwt ~ (head_cm + length_cm + babysex)^3, data = df))
  ) |> 
  mutate(
    rmse_main_effect = map2_dbl(main_effect_mod, test, \(mod, df) 
      sqrt(mean((df$bwt - predict(mod, newdata = df))^2))
    ),
    rmse_interact = map2_dbl(interact_mod, test, \(mod, df) 
      sqrt(mean((df$bwt - predict(mod, newdata = df))^2))
    )
  )

cv_df_results = cv_df_fit |> 
  summarize(
    mean_rmse_main_effect = mean(rmse_main_effect),
    mean_rmse_interact = mean(rmse_interact),
    sd_rmse_main_effect = sd(rmse_main_effect),
    sd_rmse_interact = sd(rmse_interact)
  )
cv_df_results
```

